---
title: "Credit Card Acceptance Model"
output: pdf_document
---
<h3> Credit Card Acceptance Model </h3>
This project will analyze the cross-section data on the credit history for a sample of applicants for a type of credit card. I will see the influence of different variables on whether or not an individual was accepted for a credit card. This data is from cran.r-project.org and was done by William H. Greene. 

The Y value is whether or not an applicant was accepted and it holds two values: yes or no. 

A data frame containing 1,319 observations on 12 variables.
**card** Factor. Was the application for a credit card accepted? 
**reports** Number of major derogatory reports.
**age** Age in years plus twelfths of a year.
**income** Yearly income (in USD 10,000).
**share** Ratio of monthly credit card expenditure to yearly income. 
**expenditure** Average monthly credit card expenditure.
**owner** Factor. Does the individual own their home?
**selfemp** Factor. Is the individual self-employed?
**dependents** Number of dependents.
**months** Months living at current address. 
**majorcards** Number of major credit cards held. 
**active** Number of active credit accounts.

When looking at the x values provided, I believe the following can better explain the y: reports, share, majorcards, and active. I chose those four x values because intuitively, a negative impact on those four would negatively impact whether or not an individual gets a credit card. 

```{r}
library(ISLR)
data("CreditCard")

library(ggplot2)
ggplot(CreditCard, aes(x=share, y=card))+ 
  geom_point(color='blue', size = 1)
```
```{r}
ggplot(CreditCard, aes(reports))+ 
  geom_histogram()
ggplot(CreditCard, aes(x=as.factor(reports), y = card)) + 
geom_boxplot()
```
```{r}
ggplot(CreditCard, aes(majorcards))+ 
  geom_histogram()
```

```{r}
ggplot(CreditCard, aes(reports))+ 
  geom_histogram()
```
```{r}
ggplot(CreditCard, aes(x=reports, y=card))+ 
  geom_point(color='blue', size = 1) + 
  labs(y="card accepted", x="reports")
```
```{r}
ggplot(CreditCard, aes(x=share, y=card))+ 
  geom_point(color='blue', size = 1) + 
  labs(y="card accepted", x="share")
```
```{r}
ggplot(CreditCard, aes(x=majorcards, y=card))+ 
  geom_point(color='blue', size = 1) + 
  labs(y="card accepted", x="majorcards")
```

```{r}
ggplot(CreditCard, aes(x=active, y=card))+ 
  geom_point(color='blue', size = 1) + 
  labs(y="card accepted", x="active")
```
